---
title: "R dplyr exercise_01"
output: html_notebook
---

# r4ds 의 dplyr package 활용하기 복습
## Chapter 3. Data Transformation with dplyr

```{r}
#library loading
library(tidyverse)
library(dplyr)
library(nycflights13)
library(ggplot2)
library(tibble)
```

```{r}
#check data
flights
```

## dplyr basics
* filter() : filter rows by their values
* select() : select columns by their names
* arrange() : reorder the rows
* mutate() : create new variables
* summarize() : collapse many values down to a single summary


### filter()
```{r}
#january 1st flight list
filter(flights,month==1,day==1)
filter(flights, month==11|month==12)
#filter(flights, month==11|12) -> error!
#filter(flights, month %in% c(11,12)) ->ok!
```

### arrange()
```{r}
arrange(flights,year,month,day)
#arrange(flights,desc(arr_day))
#missing values are always sorted at the end
```

### select()
```{r}
select(flights,year,month,day)
#select(flights,year:day) ==> select all columns between yeaer and day(inclusive)
#select(flights,-(year:day)) ==>select all columns except those from year to day(inclusive)
```


select()의 options
* starts_with("abc")
* ends_with("xyz")
* contains("ijk")
* mathces("(.)\\1") ==>이건 아직 뭔지 모르겠다. selects variables that match a regular expression. this one matches any variables that contain repeated characters. chapter 11
* num_range("x",1:3) matches x1,x2,x3

```{r}
rename(flights,tail_num=tailnum) # rename variables
select(flights,time_hour,air_time,everything()) #if you'd like to move variables to the start of the dataframe
```

### mutate()
```{r}
flights_sml<-select(flights,
                    year:day,ends_with("delay"),distance,air_time)
mutate(flights_sml,gain=arr_delay-dep_delay,speed=distance/air_time*60) #add variables
#if you only want to keep the new variables, use transmute(), not transmutate
transmute(flights_sml,gain=arr_delay-dep_delay,speed=distance/air_time*60)
```

useful creation functions that you can use with mutate()
==> %/% %% lead(), lag() cumsum(), cumprod(), cummin(), cummax(), cummean(), RcppRoll packages, min_rank(), row_number(), dense_rank(), percent_rank(), cume_dist(), ntle(), etc...


###summarize()
```{r}
delays<-flights%>%
  group_by(dest)%>%
  summarize(
    count=n(),
    dist=mean(distance,na.rm=TRUE),
    delay=mean(arr_delay,na.rm=TRUE)
  )%>%
  filter(count>20,dest!="HNL")

delays
```

```{r}
ggplot(data=delays,mapping=aes(x=dist,y=delay))+
  geom_point(aes(size=count),alpha=1/3)+
  geom_smooth(se=FALSE)
```

```{r}
#missing values represent cancelled flights
not_cancelled<-flights%>%
  filter(!is.na(dep_delay),!is.na(arr_delay))
not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(mean=mean(dep_delay))
```

```{r}
delays<-not_cancelled%>%
  group_by(tailnum)%>%
  summarize(
    delay=mean(arr_delay)
  )
ggplot(data=delays,mapping=aes(x=delay))+geom_freqpoly(binwidth=10)
```
```{r}
#Lahman package,data Batting
batting<-as_tibble(Lahman::Batting)
batting
```
```{r}
batters<-batting%>%
  group_by(playerID)%>%
  summarize(
    ba=sum(H,na.rm=TRUE)/sum(AB,na.rm=TRUE),
    ab=sum(AB,na.rm=TRUE)
  )
batters%>%
  filter(ab>100)%>%
  ggplot(mapping=aes(x=ab,y=ba))+
  geom_point()+
  geom_smooth(se=FALSE)

#positive correlation b/w skill and opportunities to hit the ball =>obvious
batters%>%
  arrange(desc(ba))
#if you naviely sort on desc(ba), people with teh best batting averages are lucky, not skilled. only one chance
```

Useful summary functions
mean(x), median(x), sd(x),IQR(x),mad(x)(median absolute deviation),
min(x), quantile(x,0.25), max(x)
first(x),nth(x,2), last(x)
n()
n_distinct(x) : to count the number of distinct(unique)value
sum(x>10)
mean(y==0)
```{r}
not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(
    #avearge delay
    avg_delay1=mean(arr_delay),
    #average positive delay
    avg_delay2=mean(arr_delay[arr_delay>0])
  )

not_cancelled%>%
  group_by(dest)%>%
  summarize(distance_sd=sd(distance))%>%
  arrange(desc(distance_sd))


not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(
    first=min(dep_time),
    last=max(dep_time)
  )

not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(
    first_dep=first(dep_time),
    last_dep=last(dep_time)
  )

not_cancelled%>%
  group_by(year,month,day)%>%
  mutate(r=min_rank(desc(dep_time)))%>%
  filter(r %in% range(r))

not_cancelled%>%
  group_by(dest)%>%
  summarize(carriers=n_distinct(carrier))%>%
  arrange(desc(carriers))

not_cancelled%>%
  count(dest)

not_cancelled%>%
  count(tailnum,wt=distance)

not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(n_early=sum(dep_time<500))

not_cancelled%>%
  group_by(year,month,day)%>%
  summarize(hour_perc=mean(arr_delay>60))

```


Grouping by multiple variables
```{r}
daily<-group_by(flights,year,month,day)
(per_day<-summarize(daily,flights=n()))
(per_month<-summarize(per_day,flights=sum(flights)))
(per_year<-summarize(per_month,flights=sum(flights)))
```

```{r}
daily %>%
  ungroup()%>%
  summarize(flights=n())
  
```

Grouped mutates(and filters)
```{r}
flights_sml%>%
  group_by(year,month,day)%>%
  filter(rank(desc(arr_delay))<10)

popular_dests<-flights%>%
  group_by(dest)%>%
  filter(n()>365)
popular_dests

popular_dests%>%
  filter(arr_delay>0)%>%
  mutate(prop_delay=arr_delay/sum(arr_delay))%>%
  select(year:day,dest,arr_delay,prop_delay)

```



```{r}

diamonds
```
A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:

A data frame with 53940 rows and 10 variables:

price
price in US dollars (\$326–\$18,823)

carat
weight of the diamond (0.2–5.01)

cut
quality of the cut (Fair, Good, Very Good, Premium, Ideal)

color
diamond colour, from J (worst) to D (best)

clarity
a measurement of how clear the diamond is (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best))

x
length in mm (0–10.74)

y
width in mm (0–58.9)

z
depth in mm (0–31.8)

depth
total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)

table
width of top of diamond relative to widest point (43–95)
 

